---
title: "NBA Manipulation"
author: "Sam Eckhardt and Joey Miranda"
date: "Tuesday, October 21, 2014"
output: pdf_document
---

```{r}

library(ggplot2)
library(dplyr)
library(gridExtra)
library(gtable)
library(ggmap)
library(reshape2)
library(scales)
```

```{r}

onteamsalary<-nbasalaries %>%
  tbl_df(.)%>%
  select(2,4:6)%>%
  filter(player_on.roster=="TRUE")%>%
  select(1,3,4)%>%
  group_by(team,season)%>%
  summarise(totalsalary=sum(salary))

ggplot(onteamsalary,aes(x=season,y=totalsalary,fill=team,group=team))+geom_bar(stat="identity")+facet_wrap(~team,ncol=6)+theme(legend.position="none")+scale_y_continuous("Team Salary",labels = comma)
  
```

```{r}
salary=nbasalaries %>%
  tbl_df(.)%>%
  select(2:6)%>%
  filter(player_on.roster=="TRUE")%>%
  filter(season=="2014-2015")%>%
  select(1,2,5)

ggplot(salary, aes(x=reorder(player,-salary),y=salary,group=team,fill=team,head(3)))+geom_bar(stat="identity")+facet_wrap(~team,ncol=6,scales="free")+theme(legend.position="none")+scale_y_continuous("Team Salary",labels = comma)


top3 = data.frame(matrix(vector(), 0, 3, dimnames=list(c(), c("player", "team", "salary"))))
for (i in 1:30){
  tmp= salary %>%
    filter(team==levels(salary$team)[i])%>%
    arrange(desc(salary))%>%
    head(3)
  top3=rbind(top3,tmp)
}

ggplot(top3,aes(x=reorder(player,-salary),y=salary,group=team,fill=team))+geom_bar(stat="identity")+facet_wrap(~team,ncol=6,scales="free")+theme(legend.position="none")+scale_y_continuous("Team Salary",labels = comma)

```
  
## use the draft combine information and come up with something for the top 5 picks in the draft over time. 
  
```{r}
Round1=NBA.Draft.History%>%
  filter(Round=="1")%>%
  select(1,3,5:9,11:13)

Round2=NBA.Draft.History%>%
  filter(Round=="2")%>%
  select(1,3,5:9,11:13)
  
avgheight1<-Round1%>%
  select(3,8)%>%
  filter(Height_Inches!="NA")%>%
  group_by(Pick)%>%
  summarise(Mean=mean(Height_Inches))

ggplot(avgheight1, aes(Pick,Mean,color=Pick))+geom_point(size=5)+ggtitle("Average Height of Each First Round Draft Pick")


  
#80 inches is 6'8, so the #1 pick is on average over 6'9 inches, or on average a power forward or center, based on height

pick1height<-Round1%>%
  select(1,3,6,8)%>%
  filter(Height_Inches!="NA")%>%
  filter(Pick==1)

ggplot(pick1height, aes(Draft_Year,Height_Inches,group=Pick))+geom_point(size=5)+geom_line()+ggtitle("Height of First Round Draft Picks 1957-2013")
                                                                                            
pick1to6height<-Round1%>%
  select(1,3,6,8)%>%
  filter(Height_Inches!="NA")%>%
  filter(Pick<=6)

ggplot(pick1to6height, aes(Draft_Year,Height_Inches,group=Pick,color=Pick))+geom_point(size=5)+geom_line()+facet_wrap(~Pick, ncol=2,scales="free")+geom_errorbar(stat = "hline", yintercept = "mean", width=0.8,aes(ymax=..y..,ymin=..y..))+ggtitle("Height of Top Draft Picks Per Year With Average Height")



college<-Round1%>%
  select(10)
Collegeplayercount<-as.data.frame(table(college))%>%
  arrange(desc(Freq))%>%
  head(20)

ggplot(Collegeplayercount,aes(x=reorder(college,-Freq),y=Freq,fill=college))+geom_bar(stat="identity")+theme(legend.position="none")+geom_text(aes(label=Freq))+ggtitle("Top 20 Schools First Round Players Went To")
##91 players did not go to college  

  
  